Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data

### 2022-10-21 paper/manuscript.Rmd
big update
added CI options for percentage, tweaked api, fixed many small things

---
title: "Cardiometabolic statistics for US adults: an open source web application for exploratory analysis"
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---

@@

@@

**ABSTRACT**

@@

**Background**: Data from the National Health and Nutrition Examination Survey (NHANES) program are freely available. However, analysis of these data may be time-intensive and challenging.

**Objectives**: Develop, validate, and share an open-source web application that allows for interactive, exploratory analysis of NHANES data that can be used and extended upon by others.

**Methods**: We developed the open-source web application using the R programming language and several R packages. We validated the web application by using it to reproduce results from two prior studies and a report from the Centers for Disease Control. To demonstrate the application's capabilities, we estimated characteristics of adults in the US, the prevalence of hypertension, distribution of the number of antihypertensive medication classes being used, prevalence of controlled blood pressure (BP) among those with hypertension, and the prevalence of resistant hypertension.

**Results**: The estimated mean systolic BP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020. The estimated prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%), with the estimated number of US adults with hypertension increasing from 89.8 million in 1999-2000 to 115.3 million in 2017-2020 . The age adjusted mean systolic BP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension. Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020. The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively. Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020.

**Conclusions.**: The web application developed in the current analysis may increase the usage of NHANES data for research and policy decision making, ensuring results are valid, transparent, and reproducible.

@@

The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the US population. NHANES data have been used to provide nationally representative estimates for adults living in the United States (US) in multiple studies of cardiometabolic outcomes. For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels used to define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline on the proportion and number of US adults with hypertension, recommended antihypertensive medication and with uncontrolled BP.  Additionally, NHANES data have been used to track the proportion of US adults with controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.

NHANES data are publicly available and accessible through the CDC website. However, analyzing NHANES data may be challenging for a number of reasons. In particular, specific statistical techniques are required to account for the multi-stage sampling design of NHANES data and produce nationally representative estimates. Also, each NHANES cycle includes over 50 data files, each of which has detailed informational material. NHANES data collection protocols have changed over time for certain variables, and these differences should be accounted for when harmonizing or pooling data from multiple cycles. Another potential challenge is implementation of recommended standards from the CDC, such as their guidelines for suppression of unreliable results.

To address the challenges of analyzing NHANES data, we developed an open-source web-based application that allows users to obtain nationally representative summaries of cardiometabolic outcomes for US adults using NHANES data. The goal of the application is to increase the usage of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible. The application is designed to be extended with the addition of "modules", which we define as a set of variables measured among a sub-population of NHANES participants. In this manuscript, we review the design and development of the application, present nationally representative summaries of outcomes from the "BP and Hypertension module", and describe online tutorials and resources that may help users engage with the application.

**METHODS**

@@

NHANES data include in-home interviews and study examinations conducted at mobile examination centers. The interview includes questions about demographics, health behaviors, prior diagnoses, medication use, and medical history. During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded. During the study examination, height and weight were measured, blood samples were used to measure cholesterol, glycated hemoglobin and serum creatinine, and spot urine was used to measure albumin and creatinine and to conduct a pregnancy test. Systolic and diastolic BP (SBP and DBP, respectively) were measured three times. BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020.

*Features of the Open-Source Web-Based Application*

Full summaries of this application's features and associated tutorials are available. Briefly, users can select NHANES cycles to be analyzed. Contiguous cycles can be combined to increase precision. Estimates are weighted to represent the non-institutionalized US population and may incorporate age-adjustment using direct standardization. Count estimates are calibrated to match the estimated total number of US adults. Users may visualize summaries of outcomes in groups (i.e., prevalence of hypertension by sex) with all groups in the same figure or with one figure per group, and may additionally specify subsets to exclude from their analyses. All outputs created with the web application can be downloaded and saved. Following CDC recommendations, unreliable statistical estimates are automatically suppressed.

*Design of the Open-Source Web-Based Application*

The web application was created using Shiny, an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface. The "cardioStatsUSA" R package provides additional details on the web application's design and comprehensive documentation of its components.

*The Blood Pressure and Hypertension Module*

We define a module as a subpopulation of NHANES participants paired with a collection of variables. Each variable in the collection has designated roles. For example, hypertension can be an outcome and can also be used to define groups when analyzing another outcome. This module definition is general enough to allow for diverse extensions of the web application, but may not be specific enough to illustrate how the web application works. Therefore, in the current analysis, we define and present results from the "BP and hypertension" module, defined as follows.

**Subpopulation:** Beginning with 107,622 US individuals who participated in NHANES 1999-2000 to 2017- March 2020, we restricted the subpopulation to adults ≥ 18 years of age. This exclusion was applied because statistics for hypertension and BP levels in children and adolescents are markedly different than for adults. We further restricted the subpopulation to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use. After these exclusions were applied, the sub-population included data from 56,035 participants (Figure S1).

**Variables** The primary variables from the BP and hypertension module are listed in Table 1, with full definitions provided in Supplemental Table 1. Briefly, mean SBP and DBP levels were computed over all available measurements for each participant. Oscillometric BP values were calibrated to the mercury device as described previously. Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline.

*Validation of the web application*

We validated the web application by using it in combination with our BP and hypertension module to reproduce results from two prior studies and one CDC report. A full description of the validation procedure is available.

*Statistical analysis*

We conducted a series of statistical analyses using the BP and hypertension module defined above. We estimated characteristics of the population, crude and age-adjusted mean SBP, and prevalence of hypertension for US adults by NHANES cycle. To demonstrate computation of results by group, we computed the age-adjusted prevalence of hypertension by NHANES cycle for individuals with versus without chronic kidney disease. For age adjustment in these analyses, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively). We also estimated the number of US adults with hypertension by NHANES cycle from 1999-2000 through 2017-2020.

The age-adjusted distribution of the number of antihypertensive medication classes being taken was estimated for US adults with hypertension in each NHANES cycle, overall and among those who self-reported taking antihypertensive medication and had at least one antihypertensive drug class identified in their medication inventory. The age-adjusted proportion of US adults with controlled BP (i.e., SBP < 130 mm Hg and DBP < 80 mm Hg) was estimated by NHANES cycle among US adults with hypertension and who self-reported taking antihypertensive medication. The age-adjusted prevalence of resistant hypertension was estimated by NHANES cycle for US adults with hypertension taking antihypertensive medication and among those with hypertension taking ≥ 3 classes of antihypertensive medication. For age adjustment in these analyses, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively. We estimated the prevalence of hypertension in NHANES 2017-2020 by age group, sex, and race-ethnicity. Using the data pooling option in the application, we grouped the NHANES 2011-2020 cycles to estimate the proportion of pregnant women who had uncontrolled BP (i.e., SBP ≥ 130 mm Hg or DBP ≥ 80 mm Hg).

**RESULTS**

The estimated crude mean SBP for all US adults was 123 mm Hg in 1999-2000, 120 mm Hg in 2009-2010, and 123 mm Hg in 2017-2020 (Figure 1; panel A). The estimated age-adjusted mean SBP was 124, 120, and 122 mm Hg in 1999-2000, 2009-2010 and 2017-2020, respectively (Figure 1; panel B). The estimated crude prevalence of hypertension among US adults was lowest in 2009-2010 (43.3%) and highest in 2017-2020 (46.6%) (Figure 2; panel A). After age-adjustment, the prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 and 2013-2014 (43.0%) and was 44.7% in 2017-2020. (Figure 2; panel B). The number of US adults with hypertension increased from 89.8 million in 1999-2000 to 115.3 million US adults in 2017-2020 (Figure 3). The age adjusted mean SBP was 140 in 1999-2000, 134 in 2009-2010, and 136 in 2017-2020 among US adults with hypertension and 117 in 1999-2000, 115 in 2009-2010, and 117 in 2017-2020 among US adults without hypertension (Figure S2). In each NHANES cycle, the age-adjusted estimated mean SBP was 18 to 23 mm Hg higher among US adults with versus without hypertension.  In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (Table 1). In each NHANES cycle, the estimated age-adjusted prevalence of hypertension among US adults with CKD was higher versus their counterparts without CKD (Figure S3).

Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2020 compared with 44.6% in 2017-2020. (Figure 4; panel A). Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020 (Figure 4; panel B). The age-adjusted estimated prevalence of BP control among US adults with hypertension was 10.2%, 25.8%, and 23.8% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel A). For adults who self-report taking antihypertensive medication, 26.2%, 44.1%, and 43.3% had controlled BP in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure 5; panel B).

Among US adults with hypertension who self-reported taking antihypertensive medication, the prevalence of resistant hypertension was 13.7% in 1999-2000, 17.7% in 2009-2010, and 16.2% in 2017-2020 (Figure S4; panel A). Among those who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was 73.5%, 64.1%, and 67.9% in 1999-2000, 2009-2010, and 2017-2020, respectively (Figure S4; panel B).

In 2017-2020, the prevalence of hypertension was higher at progressively older age (Figure S5),  among men versus women (Figure S6), and among non-Hispanic Black adults versus non-Hispanic White adults, non-Hispanic Asian adults, Hispanic adults and adults from other race/ethnic groups (Figure S7).

**DISCUSSION**

In the current manuscript, we describe the development of an open-access web-based application that allows the calculation of nationally representative estimates for a number of BP-related outcomes using publicly available data from NHANES. Using this app, we generated statistics for some of these BP-related outcomes over time, both crude and adjusted by age. We also generated statistics stratified by characteristics of US adults (e.g., CKD), and by pooling multiple NHANES cycles to obtain more precise estimates when working with small group sizes. Researchers, stakeholders, and the public in general interested in BP-related outcomes can now use the current app to accomplish a variety of purposes.

NHANES is a program specifically designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults. Therefore, NHANES represents the ideal data source to obtain valid and reliable statistic on BP-related outcomes among US adults. Although NHANES data are publicly available, working with these data is not exempt of challenges. Examples of challenges that may face working with NHANES data include the need to download multiple files, even for a single NHANES cycle; combining multiple NHANES variables to create the measurement of interest, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data. NHANES has been conducted over many cycles and some of the methods have been updated over time. For example, NHANES switched from using sphygmomanometer devices to oscillometric devices to measure BP in 2017-2018. NHANES data users need to be careful when analyzing data across multiple cycles. Reducing barriers to using NHANES data will facilitate the use of these data to inform BP-related public health decisions and future research.

Several design decisions have been incorporated into the app.  We required participants to have only a single SBP and DBP to be included. This approach is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682). However, mean BP and the prevalence of high BP would be lower if we required multiple BP measurements (cite pmid 23126346). Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants within one SBP and DBP measurement had three SBP and DBP measurements.  We re-calibrated the NHANES weights for the calculation of population counts (cite pmid 12500213). When estimating population counts, participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes.  Weights were not recalibrated when estimating portions as participants missing data are removed from the numerator and denominator.  Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline.  We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes.  While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist.  Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach.  We sought to make decisions that would be acceptable to people who use the app and transparent to ensure the results could be described accurately.

We were able to replicate several prior manuscripts using the app.  However, some results could not be replicated.  For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014.  When estimated by the app, the prevalence of resistant hypertension was 17.7%, 16.5% and 17.2% in 2009-2010, 2011-2012, and 2013-2014, respectively (cite pmid 30580690).  The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes. For example, the prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist.  Additionally, Lexicon Plus®includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline.  The differences in results between the app and this previously published manuscript emphasize two important points. First, while we encourage users to check results when using the app with prior publications, it may not be possible to always generate results that are identical to prior publications. Second, users should be aware of the choices made in defining variables.

The web application has a number of complementary features that allow users to complete various analysis tasks. In particular, users can generate descriptive summaries of the US population spanning over 20 years, including summaries of characteristics and cardiometabolic outcomes. The flexibility of inclusion criteria paired with the large set of outcomes and stratifying variables available allow an enormous number of hypotheses to be explored. Outputs from the app can be included in scientific proposals, presentations, or articles, citing the application's publicly available documentation to contextualize results. With the ability to download summary data from the application, users can visualize or tabulate results with aesthetics and software of their choosing. The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.

With an open-source software license and publicly available documentation and data, the web application we developed is designed to be extended. The Shiny interface and R codes supporting this application can support general sub-populations of NHANES participants, such as a sub-population of participants with cholesterol data or a sub-population of pediatric survey participants. As the current analysis demonstrates creation and analysis of a sub-population of participants with BP measurements, others can use it as a template for developing extensions of the web application that engage with new sub-populations. In addition to including new sub-populations of NHANES participants, the application may be extended to include inferential statistics and hypothesis testing, and may also incorporate new cycles of NHANES data as they become available.

@@

**Table 1**: Variables included in the blood pressure and hypertension module.

@@

ft <- ft_data %>% 
 as_flextable(hide_grouplabel = TRUE,
              col_keys = c('class', 'label')) %>% 
 set_header_labels(label = "Variables included in module") %>% 
 theme_box() %>% 
 bg(i = ~ !is.na(class), bg = 'grey80') %>%
 bg(i = indent_meds_spanner, bg = 'white') %>%
 italic(i = ~ !is.na(class)) %>%
 padding(i = ~ is.na(class), padding.left = 10) %>% 
 padding(i = indent_meds_spanner, padding.left = 10) %>% 
 padding(i = indent_meds, padding.left = 20) %>% 
 width(j = 1, width = 6) %>% 
 add_footer_lines(
  values = abbrvs_write(
   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
  )
 )

@@

**Table 2**:	Race/ethnicity distribution of the population with and without hypertension in 2017-2020

@@

nhanes_data %>% 
 nhanes_summarize(key = nhanes_key,
                  outcome_variable = 'demo_race',
                  group_variable = 'htn_accaha',
                  time_values = 'most_recent') %>%
 mutate(across(where(is.numeric), table_value)) %>% 
 filter(statistic == 'percentage') %>% 
 select(-statistic, 
        -ends_with("reason"),
        -unreliable_status,
        -review_needed) %>% 
 flextable() %>% 
 theme_box() %>% 
 width(width = 1) %>% 
 align(align = 'center', part = 'all') %>% 
 align(j = 1, align = 'center', part = 'all')

>>>>

---
title: "Hypertension statistics for US adults: an open-source web application for analysis and visualization of US NHANES data"
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
    keep_md: true
---

@@

# define age weights for standardization
age_wts_overall <- c(49.3, 33.6, 10.1, 7)
age_wts_htn <- c(26.4, 43.5, 17.0, 13.2)

@@

# ABSTRACT

@@

**Background**: Data from the US National Health and Nutrition Examination Survey (NHANES) are freely available and can be analyzed to produce hypertension statistics for the non-institutionalized US population. However, the analysis of these data requires statistical programming knowledge.

**Methods**: We developed and validated a web-based application that provides hypertension statistics for US adults using 10 cycles of NHANES data, 1999-2000 through 2017-2020 without requiring users to conduct statistical programming. The application’s interface allows users to estimate crude and age-adjusted means, quantiles, and proportions. Population counts can also be estimated. To demonstrate the application’s capabilities, we estimated hypertension statistics for US adults.

**Results**: The estimated mean systolic blood pressure (BP) declined from 123 mmHg in 1999-2000 to 120 mmHg in 2009-2010 and increased to 123 mmHg in 2017-2020. The age-adjusted prevalence of hypertension (i.e., systolic BP ≥130 mmHg, diastolic BP ≥80 mmHg or self-reported antihypertensive medication use) was 47.9% in 1999-2000, 43.0% in 2009-2010, and 44.7% in 2017-2020. In 2017-2020, 115.3 million US adults had hypertension. The age-adjusted prevalence of controlled BP, defined according to the 2017 American College of Cardiology/American Heart Association BP guideline, among US adults with hypertension was 9.6% in 1999-2000, 24.6% in 2013-2014, and 21.7% in 2017-2020. After age-adjustment and among US adults who self-report taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively. The application was validated by replicating statistics from three publications.

**Conclusions**: The application developed and validated is publicly available and produced valid, transparent, and reproducible results.

@@

The National Health and Nutrition Examination Survey (NHANES) is a program conducted by the US National Center for Health Statistics (NCHS) of the Centers for Disease Control and Prevention (CDC) and is designed to assess the health and nutritional status of the non-institutionalized US population. NHANES data have been analyzed to provide hypertension statistics for non-institutionalized US adults with important policy and public health implications. For example, NHANES data have been used to estimate the impact of the lower blood pressure (BP) levels that define hypertension and controlled BP in the 2017 American College of Cardiology/American Heart Association (ACC/AHA) BP guideline versus the Seventh Joint National Committee (JNC7) BP guideline. Additionally, NHANES data have been used to track the proportion of US adults with hypertension that have controlled BP, which were included in a Call-to-Action to Control BP from the US Surgeon General.

NHANES data are publicly available and accessible through the CDC website. However, analyzing NHANES data may be challenging for a number of reasons. Specific statistical techniques are required to account for the multi-stage sampling design used to select NHANES participants and analyses need to be weighted to produce nationally representative estimates. Each NHANES cycle includes over 50 data files, each of which has detailed informational material. NHANES data collection protocols have changed over time for some variables, and these differences should be taken into account when comparing or aggregating data from multiple cycles. Also, tests need to be performed to ensure the statistical estimates are reliable, and unstable estimates should be suppressed.

To address these challenges, we developed an open-source web-based application that provides nationally representative BP and hypertension statistics for non-institutionalized US adults using NHANES data without requiring users to conduct statistical programming. The goal of the application is to increase the use of NHANES data for research and policy decision making with a focus on ensuring results are valid, transparent, and reproducible. In this manuscript, we review the design, development and validation of the application, present BP and hypertension statistics for US adults, and describe online tutorials and resources that may help users engage with the application.

# METHODS

@@

NHANES data are collected through in-home interviews and study examinations conducted at mobile examination centers. The interview included questions about demographics, health behaviors, medical history, and medication use. During the interview, the labels of medications that participants reported taking in the preceding 30 days were recorded. During the study examination, height, weight and BP were measured and blood and spot urine samples were collected. Of relevance to the application, blood samples were used to measure total and HDL cholesterol, glycated hemoglobin and serum creatinine, and the urine sample was used to measure albumin and creatinine and to conduct a pregnancy test. The protocol for the BP measurement is available online. In brief, systolic and diastolic BP (SBP and DBP, respectively) were measured three times by trained and certified physicians. BP was measured using a mercury sphygmomanometer from 1999-2000 through 2015-2016 and using an oscillometric device in 2017-2020. The mean SBP and DBP levels were computed over all available measurements for each participant. The oscillometric SBP and DBP values were calibrated to the mercury device. Antihypertensive medication classes were defined using recommendations from the 2017 ACC/AHA BP guideline. For the current analysis, we defined hypertension, BP control, and resistant hypertension according to the 2017 ACC/AHA BP guideline.  The application also has these variables defined according to the Seventh Report of the Joint National Committee on Prevention, Detection, Evaluation, and Treatment of High Blood Pressure definitions. A complete list of the variables available in the data set is provided in Table 1 with full definitions provided in Supplemental Table 1.

There were 107,622 NHANES participants in the 10 cycles from 1999-2000 to 2017-2020. We restricted the dataset to adults ≥ 18 years of age. This exclusion was applied because statistics for BP levels and hypertension in children and adolescents are markedly different than for adults. We further restricted the population to participants who completed the in-home interview and study examination, with one or more SBP and DBP measurement, and who had data on self-reported antihypertensive medication use. After these exclusions were applied, the population in this application included 56,035 participants (Figure S1).

## Features of the Open-Source Web Application

A full summary of this application’s features and associated tutorials are available online. Briefly, users can select NHANES cycles from 1999-2000 to 2017-2020 to be analyzed. Estimates are weighted to represent the non-institutionalized US population and users may incorporate age-adjustment through direct standardization. Users can restrict analyses to subsets of participants (e.g., participants who self-reported taking antihypertensive medication). When population count estimates are requested, survey weights are calibrated to account for participants missing information on SBP, DBP or self-reported antihypertensive medication use. Users may tabulate or visualize summaries, and can present results for the overall population or in subgroups defined by a stratifying variable. All figures and datasets created with the web application can be downloaded and saved. Following CDC recommendations, unreliable statistical estimates are automatically suppressed. To increase precision and reliability of estimates, contiguous NHANES cycles can be combined.

## Development and validation of the Open-Source Web Application

The web application was created using Shiny, an open-source software package that translates code from the R programming language into HTML, CSS, or JavaScript commands that create a website interface. We created the “cardioStatsUSA” R package to provide additional details on the web application’s design and comprehensive documentation of its components. We validated the web application by using it to reproduce statistics reported in two prior studies and one CDC report.

## Statistical analysis

We performed statistical analyses to demonstrate core features of the application. We estimated the mean SBP for US adults by NHANES cycle, 1999-2000 through 2017-2020, with points showing the estimated means and error bars showing the 95% confidence intervals. We made bar charts presenting the age-adjusted prevalence of hypertension and the estimated number of US adults with hypertension. For age adjustment, we used the estimated age distribution of US adults from 1999 to 2020 as the standard (49.3%, 33.6%, 10.1% and 7.0% being 18 to 44, 45 to 64, 65 to 74 and ≥75 years of age, respectively). To illustrate how results from the application appear when saved in various formats (e.g., comma separated, portable document format, Microsoft Excel), we tabulated the estimated race/ethnicity distribution of US adults with and without hypertension, separately. We demonstrated stratification by estimating the prevalence of hypertension by NHANES cycle for US adults with and without chronic kidney disease (CKD). We showed the application’s ability to suppress output when statistical estimates are unstable by attempting to estimate the distribution of BP categories (SBP/DBP < 120/80 mm Hg, 120-129/<80 mm Hg, 130-139/80-89 mm Hg, 140-159/90-99 mm Hg and ≥ 160/100 mm Hg) among pregnant women in 2017-2020. We then showed that reliable estimates can be obtained for the distribution of BP categories among pregnant women by pooling NHANES cycles from 2009-2010 through 2017-2020.

We also illustrated how core features of the application can be combined to perform customized analyses. Specifically, we estimated the age-adjusted proportion of US adults with controlled BP by NHANES cycle among non-pregnant US adults with hypertension, overall and among those who self-reported taking antihypertensive medication. We also estimated the age-adjusted prevalence of resistant hypertension by NHANES cycle for US adults with hypertension taking antihypertensive medication and among US adults with hypertension taking ≥ 3 classes of antihypertensive medication. For age adjustment in the analysis of resistant hypertension, we set the age distribution for the standard population to represent US adults with hypertension from 1999 to 2020: 26.4%, 43.4%, 17.0% and 13.2% being 18 to 44 years, 45 to 64, 65 to 74, and ≥75 years of age, respectively.

# RESULTS

The web application successfully replicated results from prior studies. Full details on how to use the application to replicate these studies are available online. Following its validation, version 0.0.1 of the application was released and deployed in a publicly available server on **DATE TBD**.(add a citation to version 0.0.1 of the cardioStatsUSA package and a citation to the url where the shiny app lives)

## Demonstration of core features

Among non-institutionalized US adults ≥ 18 years of age, the estimated mean SBP was XXX mm Hg (95% CI ) in 1999-2000, XXX mm Hg (95% CI) in 2009-2010 and XXX mm Hg (95% CI ) in 2017-2020 (**Figure 1**).  The age-adjusted prevalence of hypertension was highest in 1999-2000 (47.9%), lowest in 2009-2010 (43.0%), and 44.7% in 2017-2020 (**Figure 2**). In 1999-2000, there were an estimated 89.8 million US adults with hypertension (**Figure 3**). The number of US adults with hypertension increased to 115.3 million in 2017-2020.  In each NHANES cycle, the estimated age-adjusted prevalence of hypertension was higher among US adults with versus without CKD (**Figure S2**). In 2017-2020, a higher percentage of US adults with versus without hypertension were non-Hispanic Black (13.5% versus 9.3%) while a lower percentage of US adults with versus without hypertension were Hispanic (12.3% versus 18.1%) (**Table 2**). Among pregnant women, the distribution of BP categories could not be estimated reliably in 2017-2020 (**Figure S3; Panel A**), but was stable after pooling NHANES cycles from 2009-2010 through 2017-2020 (**Figure S3; Panel B**)

## Demonstration of customized analyses

The age-adjusted prevalence of BP control among US adults with hypertension was lowest in 1999-2000 (9.6%), highest in 2013-2014 (24.6%), and 21.7% in 2017-2020 (**Figure 4; panel A**). Among US adults who self-reported taking antihypertensive medication, 27.4%, 48.5%, and 42.8% had controlled BP in 1999-2000, 2013-2014, and 2017-2020, respectively (**Figure 4; panel B**). The age-adjusted prevalence of resistant hypertension among US adults who self-reported taking antihypertensive medication was lowest in 1999-2000 (13.7%), highest in 2005-2006 (20.4), and 16.2% in 2017-2020 (**Figure S4; panel A**). Among US adults who were taking three or more classes of antihypertensive medication, the prevalence of resistant hypertension was lowest in 2009-2010 (64.1%), highest in 2003-2004 (78.3%), and 67.9% in 2017-2020 (**Figure S4; panel B**).

<!-- Among US adults with hypertension, 61.5% were not taking antihypertensive medication in 1999-2000 compared with 44.6% not taking antihypertensive medication in 2017-2020. (Figure 4; panel A). Among US adults who self-reported taking antihypertensive medication, the percentage taking four or more classes of antihypertensive medication increased from 3.5% in 1999-2000 to 7.6% in 2017-2020 (Figure 4; panel B).  -->

# DISCUSSION

In the current manuscript, we present a web application that allows the calculation of nationally representative estimates for BP, hypertension and antihypertensive medication-related outcomes using publicly available data from NHANES. Using this application, we generated crude and age-adjusted BP and hypertension statistics. Also, we generated statistics stratified by characteristics of US adults (e.g., CKD), and pooling multiple NHANES cycles to obtain more precise estimates when working with the small sub-group of pregnant women. Researchers, clinicians, and the general public can use the application to generate customized BP and hypertension statistics for US adults.

NHANES was designed to obtain nationally representative estimates of the health and nutrition status of non-institutionalized US adults and is an ideal data source to obtain statistics related to hypertension. SBP and DBP were measured following a standardized protocol by trained and certified staff.  NHANES data are publicly available to download but working with these data requires understanding variable definitions and advanced programming and statistical knowledge. Challenges that users encounter when analyzing with NHANES include the need to download and merge multiple data files, even for a single NHANES cycle; combining multiple variables to create outcome definitions, which may require dealing with missing data and questionnaire skip patterns; harmonizing definitions across multiple NHANES cycles; and the analyses of complex survey design data. The web application that we present in the current manuscript was designed to reduce barriers to using NHANES so that these data can more fully inform public health decisions and future research related to hypertension.

Several design decisions have been incorporated into the application. We required participants to have a single SBP and DBP to be included while some prior analyses required three SBP and DBP measurements.  We chose an approach that is consistent with several analyses conducted by CDC investigators (cite pmid 26633197, 29155682). However, mean BP and the prevalence of high BP would be lower if we required multiple BP measurements (cite pmid 23126346). Any bias resulting from this decision is likely to be small as over 95% of adult NHANES participants with at least one SBP and DBP measurement had three SBP and DBP measurements. The application re-calibrates the NHANES weights for the estimation of population counts (cite pmid 12500213). When estimating population counts, participants missing data on SBP, DBP or antihypertensive medication use cannot have BP or hypertension-related outcomes. Weights were not re-calibrated when estimating proportions as participants missing data are removed from the numerator and denominator. Medication classes were coded using generic names and the drug classes in the 2017 ACC/AHA BP guideline. We recognize the NCHS recommends using Lexicon Plus®, a proprietary database, to categorize medication classes. While the categorization of most medications is identical using generic drug names and Lexicon Plus®, some differences exist. Many additional decisions were made regarding the definitions of variables, inclusion of study participants, and analytic approach. We sought to make decisions that would be widely acceptable and transparent to ensure the results could be described accurately.

We were able to replicate several prior manuscripts using the application. However, we were not able to replicate some results. For example, a prior manuscript reported the prevalence of resistant hypertension defined by the 2017 ACC/AHA BP guideline to be 19.7% in 2009-2014(cite pmid 30580690). When estimated by the application, the prevalence of resistant hypertension over this time period was X%. The difference in the prevalence estimates can be attributed to the approach used to categorize medication into classes. The prior publication used Lexicon Plus®, which counted spironolactone as two drug classes, a potassium-sparing diuretic and an aldosterone antagonist. We included spironolactone as a single drug class, an aldosterone antagonist. Additionally, Lexicon Plus® includes Sotalol as a beta blocker and nitroglycerine as a direct vasodilator and we did not include these drugs as antihypertensive medications as they are not listed in the 2017 ACC/AHA BP guideline. The differences in results between the application and this previously published manuscript emphasize that it may not be possible to always generate results that are identical to prior publications. Users should be aware of the choices made in defining variables as this may affect the statistical estimates generated using the application.

The application has a number of features that can be used in combination to create highly customized statistics. Variables can be analyzed as outcomes or used to stratify results or restrict analyses to subsets of participants.  Combining the variables, tens of thousands of unique statistics can be estimated. In addition, users can generate statistics for US adults over 22 years, pooling results or stratifying the data. All results from the application can be saved as an image and included in scientific proposals or presentations. In addition, the ability to download summary data from the application allows users to further customize their results in tabular or graphical formats. The web application can also be used for educational purposes, as teachers can use it to analyze health outcomes for US adults interactively with students.

With an open-source software license and publicly available documentation and data, the web application we developed can be extended. The current manuscript described a set of hypertension variables and a sub-population of NHANES participants with data on BP and antihypertensive medication use. The application’s interface and its supporting R package allows additional variables to be added. In addition, the application may be extended to include data from additional NHANES cycles as they become publically available.

## Strengths/limitations

This study has a number of strengths. We used NHANES data, which are publicly available, rigorously collected, and allow for estimation of nationally representative statistics. Also, we leveraged open-source software to ensure that the application developed in the current study is transparent, able to be extended, and freely available. In addition to the application, we developed online resources with tutorials for new users and extensive documentation to contextualize results. This study also has several limitations. First, NHANES participants had their mean BP measured during a single visit, and BP guidelines recommend obtaining mean BP using at least two BP values measured on separate days. Second, the response rate for NHANES has declined from 1999-2000 through 2017-2020. Effects of this decline are unclear.

## Conclusions

We developed a web-based application for analysis of hypertension outcomes among non-institutionalized adults living in the US from 1999-2000 through 2017-2020. The application is publicly available and produces valid, transparent, and reproducible results.

@@

### Table 1: Blood pressure, hypertension and antihypertensive medication variables that may be analyzed using the web application.

@@

ft <- ft_data %>% 
 as_flextable(hide_grouplabel = TRUE,
              col_keys = c('class', 'label')) %>% 
 set_header_labels(label = "Variable") %>% 
 theme_box() %>% 
 bg(i = ~ !is.na(class), bg = 'grey80') %>%
 bg(i = indent_meds_spanner, bg = 'white') %>%
 italic(i = ~ !is.na(class)) %>%
 padding(i = ~ is.na(class), padding.left = 10) %>% 
 padding(i = indent_meds_spanner, padding.left = 10) %>% 
 padding(i = indent_meds, padding.left = 20) %>% 
 width(j = 1, width = 6) %>% 
 add_footer_lines(
  values = abbrvs_write(
   abbrvs[c('BP', 'SBP', 'DBP', 'ACC', 'AHA', 'JNC7')]
  )
 )

@@

### Table 2:	Race/ethnicity distribution of US adults with and without hypertension in 2017-2020

@@

rspec <- round_spec() %>% 
 round_using_decimal(digits = 1)
